---
title: "STAT101 Final Project: "
author: "Yijun Zhou"
date: "2025-03-05"
output: html_document
---

```{r setup, include=FALSE}
# a setup chunk that won't be included in the final document
knitr::opts_chunk$set(echo = TRUE) # show the R code below in the final document
```

```{r}
# prepare coding environment
library(tidyverse) # import package tidyverse
rm(list = ls()) # clean up the workspace
```

```{r}
# read in the "population" data and look at the first few rows
long_population = read.csv("population.csv", sep = ",")
head(long_population)
```
```{r}
colnames(long_population)<- c("country", "year", "population")
head(long_population)
```

```{r}
# plot the population over time
ggplot(long_population, aes(x = year, y = population, color = country)) + 
  geom_line(show.legend = FALSE) + # hide the legend since it's too big
  labs(title = "Population over time", x = "Year", y = "Population")
```
<br>
There are too many countries to plot all of them in one graph. Let's focus on the top 10 countries with the largest population in 2023.  

## First, let's check the basic information of the dataset.
```{r}
# check up the basics of the dataset
num_country = length(unique(long_population$country))
num_years = length(unique(long_population$year))
summary = summary(long_population)
summary
```

## Second, let's find the top 10 countries with the largest population in 2023.
```{r}
# rank the countries by the population size in 2023
population_2023 = long_population %>%
  filter(year == 2023) %>%
  arrange(desc(population))
head(population_2023)
```

## Third, let's remove the entries that are not referring to a specific country.
```{r}
# remove those are not referring to a specific country
# first, read-in world country names
countries_of_the_world = read.csv("countries_of_the_world.csv")
head(countries_of_the_world)

# remove spaces from the country names
countries_of_the_world$Country = gsub(" ", "", countries_of_the_world$Country)
head(countries_of_the_world)

# remove spaces flanking the region names
countries_of_the_world$Region = gsub("^\\s+|\\s+$", "", countries_of_the_world$Region)
head(countries_of_the_world)
```
```{r}
# select only the column named "Country"
world_country_names = countries_of_the_world$Country
head(world_country_names)
```
```{r}
#Remove spaces from the country names
world_country_names = gsub(" ", "", world_country_names)
head(world_country_names)
```
```{r}
#Remove spaces from column "country" in population_2023
population_2023$country = gsub(" ", "", population_2023$country)

# filter out the countries that are not in the world_country_names
country_population_2023 = population_2023 %>%
  filter(population_2023$country %in% world_country_names)
head(country_population_2023)
num_country_2023 = length(unique(country_population_2023$country))
```

## Fourth, let's plot the population of the top 10 countries with the largest population in 2023.
```{r}
# plot the population of the top 10 countries with the largest population in 2023
top_10_populations_2023 = country_population_2023 %>%
  top_n(10, population) %>%
  arrange(desc(population))
head(top_10_populations_2023)
```
```{r}
# save the top 10 countries with the largest population in 2023 to a file
write.csv(top_10_populations_2023, "top_10_populations_2023.csv", row.names = FALSE)
```
```{r}
# create a bar plot
ggplot(top_10_populations_2023, aes(x = reorder(country, -population), y = population, fill = country)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 10 countries with the largest population in 2023", x = "Country", y = "Population") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
```{r}
# extract data from "long_population" for the top 10 countries
top_10_country_names = unique(top_10_populations_2023$country) # create a list of names unique to the top 10 countries
head(top_10_country_names)
```
```{r}
# remove spaces from the country names in "long_population"
long_population$country = gsub(" ", "", long_population$country)

top_10_population = long_population %>%
  filter(country %in% top_10_country_names) # filter out the data for the top 10 countries

# plot the population over time for the top 10 countries
ggplot(top_10_population, aes(x = year, y = population, color = country)) +
  geom_line() +
  labs(title = "Population over time for the top 10 countries", x = "Year", y = "Population")
```
## Fifth, let's extract country names from "countries_of_the_world" for each region.
```{r}
# count unique entries in "Region" column in "countries_of_the_world".
unique_regions = unique(countries_of_the_world$Region)
num_region = length(unique_regions)
# list all the unique regions
unique_regions
```
```{r}
# extract names of countries from "countries_of_the_world" based on each region
countries_by_region = list() # create an empty list to store the names of countries for each region
for (i in 1:num_region) {
  region = unique_regions[i]
  countries = countries_of_the_world %>%
    filter(Region == region) %>%
    select(Country) %>%
    unlist() # extract the names of countries for the region
  countries_by_region[[region]] = countries
}

# remove spaces from the country names
for (i in 1:num_region) {
  region = unique_regions[i]
  countries = countries_by_region[[region]]
  countries = gsub(" ", "", countries)
  countries_by_region[[region]] = countries
}

# for each region, extract the data from "long_population"
population_by_region = list() # create an empty list to store the data for each region
for (i in 1:num_region) {
  region = unique_regions[i]
  countries = countries_by_region[[region]]
  population = long_population %>%
    filter(country %in% countries) # extract the data for the countries in the region
  population_by_region[[region]] = population
}

# check the data for the first region
head(population_by_region[[1]])

# save the data for each region to a file
for (i in 1:num_region) {
  region = unique_regions[i]
  population = population_by_region[[region]]
  write.csv(population, paste(region, ".csv"), row.names = FALSE)
}
```
## Sixth, let's plot the population over time for each region.
```{r}
# plot the population over time for each region and save the plots
for (i in 1:num_region) {
  region = unique_regions[i]
  population = population_by_region[[region]]
  plot = ggplot(population, aes(x = year, y = population, color = country)) +
    geom_line() +
    labs(title = paste("Population over time for", region), x = "Year", y = "Population")
  ggsave(paste(region, ".png"), plot) 
}

# make facet plots for all regions
# combine the data for all regions and add a column for the region
all_regions = bind_rows(population_by_region, .id = "Region")
ggplot(all_regions, aes(x = year, y = population, color = country)) +
  geom_line(show.legend = FALSE) +
  labs(title = "Population over time for all regions", x = "Year", y = "Population") +
  facet_wrap(~Region)
```
## Seventh, let's calculate the total population for each region in 2023.
```{r}
# calculate the total population for each region in 2023
total_population_region_2023 = list() # create an empty list to store the total population for each region
for (i in 1:num_region) {
  region = unique_regions[i]
  population = population_by_region[[region]]
  total = sum(population$population[population$year == 2023])
  total_population_region_2023[[region]] = total
}

# create a data frame for the total population in 2023
total_population_region_2023_df = data.frame(Region = unique_regions, total_population_region_2023 = unlist(total_population_region_2023))
total_population_region_2023_df

# save the total population in 2023 for each region to a file
write.csv(total_population_region_2023_df, "total_population_region_2023.csv", row.names = FALSE)
```

```{r}
# create a bar plot for the total population in 2023 for each region
ggplot(total_population_region_2023_df, aes(x = reorder(Region, -total_population_region_2023), y = total_population_region_2023, fill = Region)) +
  geom_bar(stat = "identity") +
  labs(title = "Total population in 2023 for each region", x = "Region", y = "Total Population") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Eighth, let's calculate the average population growth rate for each region.
```{r}
# calculate the average population growth rate for each region
average_growth_rate = list() # create an empty list to store the average growth rate for each region
for (i in 1:num_region) {
  region = unique_regions[i]
  population = population_by_region[[region]]
  growth_rate = (population$population[population$year == 2023] - population$population[population$year == 2000]) / 23
  average = mean(growth_rate)
  average_growth_rate[[region]] = average
}

# create a data frame for the average growth rate for each region
average_growth_rate_df = data.frame(Region = unique_regions, Average_Growth_Rate = unlist(average_growth_rate))
write.csv(average_growth_rate_df, "average_growth_rate(2000-2023).csv", row.names = FALSE)

# create a bar plot for the average growth rate for each region
ggplot(average_growth_rate_df, aes(x = reorder(Region, -Average_Growth_Rate), y = Average_Growth_Rate, fill = Region)) +
  geom_bar(stat = "identity") +
  labs(title = "Average growth rate between 2000 and 2023", x = "Region", y = "Population Per Year") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
## Ninth, let's do some statistical analysis on the population data.
```{r}
# calculate the mean and standard deviation of the population per country of each region in 2023
mean_population_2023 = list() # create an empty list to store the mean population for each region
sd_population_2023 = list() # create an empty list to store the standard deviation of the population for each region
for (i in 1:num_region) {
  region = unique_regions[i]
  population = population_by_region[[region]]
  mean = mean(population$population[population$year == 2023])
  sd = sd(population$population[population$year == 2023])
  mean_population_2023[[region]] = mean
  sd_population_2023[[region]] = sd
}

# create a data frame for the mean and standard deviation of the population in 2023 for each region
population_stats_2023_df = data.frame(Region = unique_regions, Mean_Population_2023 = unlist(mean_population_2023), SD_Population_2023 = unlist(sd_population_2023))
head(population_stats_2023_df)
```
```{r}
write.csv(population_stats_2023_df, "population_stats_region_2023.csv", row.names = FALSE)
```


```{r}
# create a boxplot of the mean population in 2023 for each region
ggplot(population_stats_2023_df, aes(x = Region, y = Mean_Population_2023, fill = Region)) +
  geom_boxplot() +
  labs(title = "Mean population in 2023 for each region", x = "Region", y = "Mean Population") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
# perform ANOVA test to compare if the mean population in 2023 is significantly different among regions
anova_test = aov(Mean_Population_2023 ~ Region, data = population_stats_2023_df)
anova_test
```
```{r}
# check the assumptions of ANOVA
qqnorm(residuals(anova_test))
```

```{r}
# perform Tukey's HSD test to compare the mean population in 2023 between each pair of regions
tukey_test = TukeyHSD(anova_test)
tukey_test
```



