---
title: "Capstone Dengue Virus Interactive Map"
output:
  html_document:
    df_print: paged
---

# 1. Read in the Dengue Virus Metadata
```{r}
setwd("/Users/yijunzhou/My Drive/Data Analysis_H2L2C/0816 Capstone_Dengue_Virus_Interactive_Map")
library(tidyverse)
combined_metadata <- as_tibble(read_tsv("/Users/yijunzhou/My Drive/Data Analysis_H2L2C/0816 Capstone_Dengue_Virus_Interactive_Map/20230727_dengue_patient_metadata_combined.tsv"))
```

# 2.Look at "combined_metadata"
```{r}
names(combined_metadata)
head(combined_metadata)
```

# 3.Extract columns "Serotype" and "Location"
```{r}
sero_loc <- combined_metadata %>% select(Serotype, Location)
head(sero_loc)
```

# 4.Visualize "sero_loc"
```{r}
ggplot(sero_loc, aes(x = Location)) + geom_bar() 
ggplot(sero_loc, aes(x = Location, fill = Serotype)) + geom_bar() 
# x axis is too cramped, too many locations to plot

ggplot(sero_loc, aes(x = Serotype, fill = Serotype)) + geom_bar()
# shows the number counts for each serotype
```

# 5.Separate "Locations" into "Continent" and "Countries"
```{r}
# how many distinct locations are there in the data frame?
sero_loc %>% distinct(Location)   #need to tidy-up the data

sero_cont_count <- separate(sero_loc,"Location", into = c("Continents","Countries"), sep = "/") %>% 
  mutate(across(where(is.character), str_trim))  #remove the extra spaces in "Continents"and "Countries"
head(sero_cont_count)
```

# 6.Serotype vs Continents
```{r}
ggplot(sero_cont_count, aes(x = Continents, fill = Serotype)) + 
  geom_bar()   
```

# 7.Serotype vs Countries
```{r, fig.height = 8, fig.width = 20}
ggplot(sero_cont_count, aes(x = Countries, fill = Serotype)) + 
  geom_bar() +              
  theme (axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title = "Number of Sequences Based on Serotypes and Countries") +
  ylab("Number of Reported Sequences") +
  xlab("Countries")
```

# 8. Serotype vs Countries vs Continents + Legends
```{r, fig.height = 10, fig.width = 20}
ggplot(sero_cont_count, aes(x = Countries, fill = Serotype), xaxt) + 
  geom_bar() +
  facet_wrap(~Continents, nrow = 2) +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  labs(title = "Number of Sequences vs Serotypes vs Countries vs Continents") +
  ylab("Number of Reported Sequences") +
  xlab("Countries") +
  theme(strip.text.x = element_text(size = 10))
```

# 9. Count the number of dengue virus data entries per country
```{r}
head(sero_cont_count)

count_dengue <- sero_cont_count %>% 
  select(Countries) %>% 
  group_by(Countries) %>% 
  count()

head(count_dengue)
```

# 10. Prepare the geospatial data
```{r}
library(rgdal) # to read geospatial data into R
library(sf) # to manipulate geospatial data

# load geospatial data into R
world_spdf <- readOGR( 
  dsn= ("/Users/yijunzhou/My Drive/Data Analysis_H2L2C/0808 Leaflet Tutorial"), 
  layer="TM_WORLD_BORDERS_SIMPL-0.3",
  verbose=FALSE
)
```

# 11. Merge the geospatial data with dengue virus counts based on country (count_dengue)
```{r}
# rename the count_dengue: n -> dengue
colnames(count_dengue)[2] <- "dengue"

# merge world_spdf@data with count_dengue using dplyr package
# world_spdf_dengue <- world_spdf@data %>%
#  left_join(. , count_dengue, by=c("NAME"="Countries")) # This is working, but the output is not spdf

# need to fix some names in world_spdf
count_dengue[131,1] <- "Viet Nam" #Vietnam
count_dengue[65,1] <- "Lao People's Democratic Republic" #Laos
count_dengue[109,1] <- "Korea, Republic of" #South Korea
             

# try merge from sp package, the first argument must be spdf
world_spdf_dengue2 <- 
  merge(world_spdf, count_dengue, by.x = "NAME", by.y = "Countries")

# Note that if the number of dengue virus is NA, it is in fact 0
world_spdf_dengue2@data$dengue[is.na(world_spdf_dengue2@data$dengue)] = 0.001
world_spdf_dengue2@data$dengue <- as.numeric(as.character(world_spdf_dengue2@data$dengue))
## use world_spdf_dengue2 to plot
```

# 12.Visualize the distribution of Dengue Virus per country
```{r}
world_spdf_dengue2@data %>% 
  ggplot( aes(x=dengue)) + 
    geom_histogram(bins=20, fill='#69b3a2', color='white') +
    xlab("Total Dengue Virus Entries") + 
    theme_bw()
```

# 13.Plot using leaflet choropleth map
```{r}
library(htmlwidgets) # to save widget in a html file
library(leaflet) # to make the interactive map

# Create a color palette with handmade bins.
library(RColorBrewer)
mybins <- c(0,5,10,50,500,1000,5000,Inf)
mypalette <- colorBin( palette="YlOrBr", domain=world_spdf_dengue2@data$dengue, na.color="transparent", bins=mybins)

# Prepare the text for tooltips:
mytext <- paste(
    "Country: ", world_spdf_dengue2@data$NAME,"<br/>", 
    "Area: ", world_spdf_dengue2@data$AREA, "<br/>", 
    "Dengue Virus Entries: ", round(world_spdf_dengue2@data$dengue, 2), 
    sep="") %>%
  lapply(htmltools::HTML)

# Final Map
m <- leaflet(world_spdf_dengue2) %>% 
  addTiles()  %>% 
  setView( lat=10, lng=0 , zoom=2) %>%
  addPolygons( 
    fillColor = ~mypalette(dengue), 
    stroke=TRUE, 
    fillOpacity = 0.9, 
    color="white", 
    weight=0.3,
    label = mytext,
    labelOptions = labelOptions( 
      style = list("font-weight" = "normal", padding = "3px 8px"), 
      textsize = "13px", 
      direction = "auto"
    )
  ) %>%
  addLegend( pal=mypalette, values=~dengue, opacity=0.9, title = "Dengue Virus Entries", position = "bottomleft" )
m

#saveWidget(m, file="/Users/yijunzhou/My Drive/Data Analysis_H2L2C/0816 Capstone_Dengue_Virus_Interactive_Map/dengue_map1.html")

```

