---
title: "Social media data analysis"
author: "Yijun Zhou"
date: "2025-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is an analysis of social media data from kaggle: https://www.kaggle.com/datasets/atharvasoundankar/viral-social-media-trends-and-engagement-analysis

As part of the STAT 101 class, I will:<br>
1. Run a statical analysis on the data. <br>
2. Make a visualization of the data. <br>
3. Record a presentation of the analysis. <br>

## First, let's read-in the data
```{r}
# read-in the dataset from the csv file "viral_social_media_trends.csv"
social_media_data <- read.csv("viral_social_media_trends.csv")

# check the first few rows of the dataset
head(social_media_data)
```

## Second, prepare the coding environment
```{r}
library(tidyverse)
library(hrbrthemes)
library(viridis)
```


## Third, make a visualization of the data to check the distribution of "Likes" per "Platform".
```{r}
# create a boxplot of "Platform" and "Likes"
social_media_data %>%
  ggplot( aes(x=Platform, y=Likes)) +
  geom_boxplot() +
  scale_y_log10() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6, option = "magma") +
  geom_jitter(color="blue", size=0.4, alpha=0.9) +
  theme_ipsum(grid = TRUE, axis = TRUE,axis_col = "black") +
  theme(
    legend.position="none",
    plot.title = element_text(size=11),
    axis.title.y = element_text(vjust = 5, hjust = 0.5)
  ) +
  ggtitle("Viral Social Media Likes Per Platform") +
  xlab("Platform") + 
  ylab("Likes")
```

## How about the distribution of "Shares" per "Likes"?
```{r}
# create a scatterplot of "Likes" and "Shares"
social_media_data %>%
  ggplot( aes(x=Likes, y=Shares)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6, option = "magma") +
  theme_ipsum(grid = TRUE, axis = TRUE,axis_col = "black") +
  theme(
    legend.position="none",
    plot.title = element_text(size=11),
    axis.title.y = element_text(vjust = 5, hjust = 0.5)
  ) +
  ggtitle("Viral Social Media Likes vs Shares") +
  xlab("Likes") + 
  ylab("Shares")
```
```{r}
# create a densityplot of "Likes" and "Shares"
social_media_data %>%
  ggplot( aes(x=Likes, fill=Shares)) +
  geom_density(alpha=0.6) +
  scale_x_log10() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6, option = "magma") +
  theme_ipsum(grid = TRUE, axis = TRUE,axis_col = "black") +
  theme(
    legend.position="none",
    plot.title = element_text(size=11),
    axis.title.y = element_text(vjust = 5, hjust = 0.5)
  ) +
  ggtitle("Viral Social Media Likes vs Shares") +
  xlab("Likes") + 
  ylab("Shares")
```


## Fourth, calculate the correlations
```{r}
cor.test(social_media_data$Likes, social_media_data$Shares, method = "spearman")
```

